---
title: "Models"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(car)
library(xtable)
library(broom)
```

## Setup

```{r prework ,message=FALSE, warning=FALSE}
alpha <- 0.05

find_sig <- function(model, alpha, bonferonni = F){
  if(bonferonni){
  model1_sum %>%
    filter(p.value < alpha/(nrow(model1_sum))) %>%
    select(term, estimate, p.value)
  }
  else{
  model1_sum %>%
    filter(p.value < alpha) %>%
    select(term, estimate, p.value)
  }
}

colleges <- read_csv(paste0("https://raw.githubusercontent.com/",
                           "ayakkala1/final-stat-334/",
                           "master/data/USNews.csv"))

colleges <- colleges %>%
              drop_na()

head(colleges)
```

## Model Information

Response: __GradRate__ (in percent of students who graduate within 6 years)

###  Variables for USNews Data

1. College name
2. State (postal code)
3. Funding (categorical variable: Public or Private)
4. Number of applications received (in number of students)
5. Number of applicants accepted (in number of students)
6. Number of new students enrolled (in number of students)
7. Pct. new students from top 10% of H.S. class (in percent)
8. Number of fulltime undergraduates (in number of students)
9. Out-of-state tuition (in dollars)
10. Room and board costs (in dollars)
11. Pct. of faculty with Ph.D.'s (in percent)
12. Student/facultyratio (instudents per faculty member)
13. Pct.alumni who donate (in percent)
14. Instructionalexpenditureperstudent (in dollars)
15. Graduation rate (in percent of students who graduate within 6 years)
16. StdScore = the average of standardized SAT and ACT scores for students at the school (in standard deviations above or below average)

Missing values are denoted with *

## Default Model

```{r Default Model Summary}
explanatory <- names(colleges)[names(colleges) != "Gradrate"]

model1 <- lm(Gradrate ~ Funding + Apps + Accepted + 
                        Enrolled + Top10 + FTUG + 
                        Tuition + RmBrd + PhD + 
                        SFRatio + Alumni + Spending + StdScore, 
             data = colleges)

model1_sum <- tidy(summary(model1))

model1_sum
```

```{r Significant Variables}
find_sig(model1_sum, alpha, bonferonni = F)
```

```{r Correlation with Response}
numerics <- unlist(lapply(colleges, is.numeric))

correlations <- c()
numeric_names <- names(colleges[,numerics])

for (i in colleges[,numerics]){
  correlations <- c(correlations, cor(colleges$Gradrate, i))
}

cor_with_grad <- as_tibble(cbind(numeric_names, correlations))

cor_with_grad %>%
  mutate(correlations = as.numeric(correlations)) %>%
  filter(numeric_names != "Gradrate") %>%
  mutate(numeric_names = fct_reorder(numeric_names, correlations)) %>%
  ggplot(aes(x = numeric_names, y = correlations, fill = numeric_names)) +
          geom_col() + coord_flip() + guides(fill = FALSE) +
          xlab(element_blank()) + ylab("Correlation") +
          ggtitle("Correlation with Graduation Rate") +
          scale_y_continuous(breaks = seq(-1, 1, by = 0.1))
```

```{r corrplot, message=FALSE, warning=FALSE, fig.height=8, fig.width = 8}
library(corrplot)

college_corr <- cor(colleges[, numerics], method = "pearson")

corrplot(college_corr, 
         type = "upper", 
         order = "hclust", 
         tl.col = "black", 
         tl.srt = 45)
```

